---
title: "accelerometerData"
author: "Ian Kim"
date: "April 25, 2015"
output: html_document
---

# Read and Subset Data
Import libraries and read training data set.
```{r}
library(caret)
library(randomForest)

data <- read.csv("pml-training.csv")
```

Subset data for columns in data set containing accelerometer data only. Omit the 'var' variables, which are 'NA' for more than 50% of the observations. Also omit the 'total' variables, which appear to have been derived from the x, y, z variables so are assumed to be correlated with these variables. 
```{r}
df <- data.frame(data[,grep('accel', names(data))], data[160])
df <- data.frame(df[,-grep('var', names(df))])
df <- data.frame(df[,-grep('total', names(df))])
```

Per below, no variables are near zero variance, so no more variables will be omitted.
```{r}
nsv <- nearZeroVar(df, saveMetrics=TRUE)
nsv
```


# Train and Evaluate Model
Fit a model using random forests. Using this method, cross-validation is done internally in the process of generating the fitted model. The out-of-sample accuracy is estimated to be the out-of-bag (OOB) estimate (reference: http://www.stat.berkeley.edu/~breiman/RandomForests/cc_home.htm#ooberr). Therefore, a separate explicit cross-validation step to get an unbiased estimate of the out-of-sample error is not required. We will fit the model on this entire data set.

```{r}
rf_model <- randomForest(classe~., data=df)
rf_pred <- predict(rf_model, df)
print(confusionMatrix(rf_pred, df$classe))
print(rf_model)
```
This model achieves 100% accuracy on the training set. The out-of-sample accuracy is estimated to be the out-of-bag (OOB) estimate of error rate printed above: 4.3%.


# Final Predictions Using Testing Set
Finally, use this model to predict classes for the outcome variable for the separate testing set provided.
```{r}
data_testing <- read.csv("pml-testing.csv")
final_pred <- predict(rf_model, data_testing)
final_pred

# write files for submission
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(final_pred)
```
